\documentclass[12pt]{article}
\usepackage[top=1.0in, bottom=0.75in, left=0.75in, right=2.5in]{geometry}
\usepackage{fancyhdr, textpos, enumerate}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{ebgaramond-maths, plex-mono}
\usepackage[makeroom]{cancel}

\pagestyle{fancy}
\fancyhf{}
\setlength\headheight{28pt}
\setlength\headwidth{7in}
\setlength{\TPHorizModule}{1pt}
\setlength{\TPVertModule}{1pt}
\lhead{\textsc{group theory\\spring 2018}}
\rhead{\textsc{homework bundle \#3\\page \thepage}}
\chead{\textsc{rob\\ireton}}

\newcommand{\zee}{\mathbb{Z}}

\newenvironment{exercise}[2]{\begin{textblock}{32}[1,0](0,#2)\noindent#1\end{textblock}}{\vspace{1in}}

\begin{document}
\begin{exercise}{1.47}{8}
	{\noindent}Let $G$ be a group and fix $a\in G$. Consider the function $\gamma_a:G\to G$ given by $\gamma_a(x)=axa^{-1}$. (This is called conjugation by $a$.)
	\begin{enumerate}[(i.)]
		\item Prove that $\gamma_a$ is an isomorphism.
		\bigskip
		\begin{proof}
			Let $x,y\in G$.\\
			\\
			$\gamma_a(xy)=axya^{-1}=ax1ya^{-1}=axa^{-1}aya^{-1}=\gamma_a(x)\gamma_a(y)$.
			Since $\gamma_a(xy)=\gamma_a(x)\gamma_a(y)$, $\gamma_a$ is a homomorphism.\\
			\\
			Suppose $\gamma_a(x)=\gamma_a(y)$. Then
			\begin{align*}
				axa^{-1}&=aya^{-1}\\
				axa^{-1}a&=aya^{-1}a\\
				a^{-1}axa^{-1}a&=a^{-1}aya^{-1}a\\
				1x1&=1y1\\
				x&=y.
			\end{align*}
			$\gamma_a$ is injective.\\
			\\
			Since $a$ and $y$ are in $G$, $a^{-1}ya$ is also in $G$. But $\gamma_a(a^{-1}ya)=aa^{-1}yaa^{-1}=y$. $\gamma_a$ is surjective.\\
			\\
			Since $\gamma_a$ is a bijective homomorphism, it is, by definition, an isomorphism.
		\end{proof}

		\item If $a,b\in G$, prove that $\gamma_a\gamma_b=\gamma_{ab}$.
		\bigskip
		\begin{proof}
			Let $x\in G$. $\gamma_{ab}(x)=(ab)x(ab)^{-1}=abxb^{-1}a^{-1}$. This is the same as $\gamma_a\gamma_b$, the composition of functions $\gamma_a$ and $\gamma_b$: $\gamma_a\circ \gamma_b(x)=a(bxb^{-1})a^{-1}$.
		\end{proof}
	\end{enumerate}
\end{exercise}

\newpage

\begin{exercise}{1.50}{26}
	\quad
	\begin{enumerate}[(i.)]
		\item Prove that a group $G$ is abelian if and only if the function $f:G\to G$ given by $f(a)=a^{-1}$ is a homomorphism.
		\bigskip
		\begin{proof}
			First, suppose that $G$ is abelian. Let $a,b\in G$; then $f(ab)=(ab)^{-1}=b^{-1}a^{-1}$. $G$ is abelian so $b^{-1}a^{-1}=a^{-1}b^{-1}$. But, $a^{-1}b^{-1}=f(a)f(b)$; so $f$ must be a homomorphism.\\
			\\
			Next, suppose that $f$ is a homomorphism and let $a,b\in G$. Then:
			\begin{align*}
				f(ab)&=f(a)f(b)\\
				(ab)^{-1}&=a^{-1}b^{-1}\\
				b^{-1}a^{-1}&=a^{-1}b^{-1}\\
				bb^{-1}a^{-1}&=ba^{-1}b^{-1}\\
				a^{-1}b&=ba^{-1}b^{-1}b\\
				a^{-1}ba&=ba^{-1}a\\
				aa^{-1}ba&=ab\\
				ba&=ab.
			\end{align*}
			So, $G$ must be abelian.
		\end{proof}

		\item Let $f:G\to G$ be an isomorphism. If $f$ has no nontrivial fixed points (\textsc{iow}, $f(x)=x\implies x=e$) and if $f\circ f$ is the identity map, prove that $\forall x\in G$, $f(x)=x^{-1}$.
		\bigskip

		Let $a,b\in G$. $f(ab)=f(a)f(b)$. $f(a)=f(b)\implies a=b$. $\forall b\in G, \exists a\in G$ such that $f(a)=b$. $f(a^{-1})=f(a)^{-1}$. $ff(a)=a$.\\
		\\
		\dots
	\end{enumerate}
\end{exercise}


\begin{exercise}{2.2}{5}
	{\noindent}If $k$ is a field, show that $\text{SL}(n,k)$, the set of all $n\times n$ matrices over $k$ having determinant $1$, is a subgroup of $\text{GL}(n,k)$.
	\bigskip
	\begin{proof}
		Let $k$ be a field and let $n\in\mathbb{N}$. $\text{GL}(n,k)$ is the group of $n\times n$ matrices with elements from $k$ having a nonzero determinant. Its identity element is the identity matrix $I_n$ and its operation is matrix multiplication.
		$\text{SL}(n,k)$ is the set of $n\times n$ matrices with elements from $k$ having a determinant equal to $1$. Observe that $\text{SL}(n,k)$ is a subset of $\text{GL}(n,k)$, specifically, all the elements of $\text{GL}(n,k)$ with determinant equal to $1$.
		We know from Linear Algebra that $\det(I_n)=1$, so $I_n\in \text{SL}(n,k)$.\\
		\\
		Let $s,t\in\text{SL}(n,k)$; then $\det(s)=1$ and $\det(t)=1$. $\det(t^{-1})=1/\det(t)=1$, so $t^{-1}\in\text{SL}(n,k)$.
		$\det(st^{-1})=\det(s)\det(t^{-1})=1$, so $st^{-1}\in\text{SL}(n,k)$.\\
		\\
		Since $\text{SL}(n,k)$ is a subset of $\text{GL}(n,k)$, $1\in \text{SL}(n,k)$ and $s, t\in \text{SL}(n,k)$ imply $st^{-1}\in\text{SL}(n,k)$, Theorem 2.2 tells us that $\text{SL}(n,k)$ is a subgroup of $\text{GL}(n,k)$.
	\end{proof}
\end{exercise}

\begin{exercise}{2.9i}{5}
	{\noindent}Prove that $S_n$ can be generated by $(1\enspace 2), (1\enspace 3),\;\dotsc, (1\enspace n)$.
	\bigskip
	\begin{proof}
		From Theorem 1.3, we know that every element of $S_n$ is a product of transpositions.\\
		\\
		For transpositions that are not of the form $(1\enspace k)$ where $1<k\leq n$, observe that $(i\enspace j)$ may be expressed as $(1\enspace i)(1\enspace j)(1\enspace i)$.\\
		\\
		So, for each transposition in the product that makes up each element of $S_n$, we have a way to express that cycle as a product of elements of the set above, giving us a way to generate all of $S_n$.
	\end{proof}
\end{exercise}

\begin{exercise}{2.10}{5}
	{\noindent}Prove: If $G$ is a finite group and $K\le H\le G$, then $[G:K]=[G:H][H:K]$.
	\bigskip
	\begin{proof}
		From Lagrange's Theorem (2.11) we have
		\begin{align*}
			[G:H]&=|G|/|H|\\
			[H:K]&=|H|/|K|\\
			[G:K]&=|G|/|K|
		\end{align*}
		All of these indexes and orders are natural numbers, so we can rearrange to get $|G|=[G:H]|H|$ and $|H|=[H:K]|K|$. Substitution gives
		\begin{align*}
			[G:K]&=|G|/|K|\\
			&=[G:H]|H|/|K|\\
			&=[G:H][H:K]|K|/|K|\\
			&=[G:H][H:K]
		\end{align*}
	\end{proof}
\end{exercise}

% \newpage
%
% \begin{exercise}{2.14}{5}
% 	{\noindent}Prove: If $a\in G$ has finite order and $f:G\to H$ is a homomorphism, then the order of $f(a)$ divides the order of $a$.
% 	\bigskip
% \end{exercise}
%
% \begin{exercise}{2.16}{5}
% 	{\noindent}Prove: If $H\le G$ has index $2$, then $\forall a\in G$, $a^2\in H$.
% 	\bigskip
% \end{exercise}

\end{document}
